Containers and R---the Rockerverse and beyond

### 2020-01-13 manuscript.Rmd
add a few more links and missing reference
update tex file

- Docker images for cloud services
  - A popular application of using Docker is to create the environments that can run in parallel for speeding up R code.  For example, [`googleComputeEngineR`](https://CRAN.R-project.org/package=googleComputeEngineR)'s `gce_vm_cluster()` function can create clusters of 2 or more virtual machines, running multi-CPU architectures.  Instead of running a local R script with the local CPU and RAM restrictions, the same code can be processed on all CPU threads of the cluster of machines in the cloud, all running in a Docker container with the same R environments.   This is achieved through `googleComputeEngineR`'s integration with the R parralisation library [`library(future)` by Henrik Bengtsson]( https://CRAN.R-project.org/package=future).  Local R computation can be thrown up to a multi-CPU and VM environment to achieve parrell computation in a few lines of R code in your local session. - [some demonstrations are available here](https://cloudyr.github.io/googleComputeEngineR/articles/massive-parallel.html). 
- [Google Cloud Run](https://cloud.run) - This is a CaaS (Containers as a Service) that lets you launch a Docker container without worrying about underlying infrastructure.  This dispenses with the developer creating a cloud server to run the Docker image on, by abstracting away those servers to a more serverless configuration.  Cloud Run lets you run your code on top of a managed or your own Kubernetes cluster running Knative, and can accept any Docker image.  The service takes care of network ingress, scaling machines up and down to zero, authentication and authorisation, all features which are non-trivial for a developer to create on their own.  This can be used to scale up R code to millions of instances if they need to, with little or no changes to existing code. An R implementation is shown here at [cloudRunR](https://github.com/MarkEdmondson1234/cloudRunR) which uses Cloud Run to create a scalable R plumber API. 
- `batchtools` \citep{Lang2017batchtools} can [schedule jobs with Docker Swarm](https://mllg.github.io/batchtools/reference/makeClusterFunctionsDocker.html)
- scalable deployments, e.g., start with numerous Shiny talks mentioning Rocker at useR!2017
- [dynmethods](https://github.com/dynverse/dynmethods) [@rcannood]: In order to evaluate ±50 computational methods which all used different environments (R, Python, C++, ...), we wrapped each of them in a docker container and can execute these methods from R. Again, all of these containers are being built on travis-ci, and will only be pushed to docker hub if the integration test succeeds.

@@

...

@@

DN conceived of the presented idea and \href{https://github.com/nuest/rockerverse-paper/issues/3}{initialised the formation of the writing team}, wrote the section about images for R distributions, wrote the conclusions section, and revised all sections.
CB ..
RC and EH wrote the section on interfaces for Docker in R.
DE wrote the introduction and section about Rocker.
ME ..
CF wrote the paragraphs about `r-online`, `dockerfiler`, `r-ci` and `r-db`.
BW ..
KR ..
NR ..
NX wrote the section on liftr.
LS \& NT wrote the section on Bioconductor.
All authors contributed minor edits, gave feedback on other sections, and approved the final version.

>>>>

- Docker images for cloud services
  - A popular application of using Docker is to create the environments that can run in parallel for speeding up R code.  For example, [`googleComputeEngineR`](https://CRAN.R-project.org/package=googleComputeEngineR)'s `gce_vm_cluster()` function can create clusters of 2 or more virtual machines, running multi-CPU architectures.  Instead of running a local R script with the local CPU and RAM restrictions, the same code can be processed on all CPU threads of the cluster of machines in the cloud, all running in a Docker container with the same R environments.   This is achieved through `googleComputeEngineR`'s integration with the R parralisation library [`library(future)` by Henrik Bengtsson]( https://CRAN.R-project.org/package=future).  Local R computation can be thrown up to a multi-CPU and VM environment to achieve parrell computation in a few lines of R code in your local session. - [some demonstrations are available here](https://cloudyr.github.io/googleComputeEngineR/articles/massive-parallel.html). 
- [Google Cloud Run](https://cloud.run) - This is a CaaS (Containers as a Service) that lets you launch a Docker container without worrying about underlying infrastructure.  This dispenses with the developer creating a cloud server to run the Docker image on, by abstracting away those servers to a more serverless configuration.  Cloud Run lets you run your code on top of a managed or your own Kubernetes cluster running Knative, and can accept any Docker image.  The service takes care of network ingress, scaling machines up and down to zero, authentication and authorisation, all features which are non-trivial for a developer to create on their own.  This can be used to scale up R code to millions of instances if they need to, with little or no changes to existing code. An R implementation is shown here at [cloudRunR](https://github.com/MarkEdmondson1234/cloudRunR) which uses Cloud Run to create a scalable R plumber API. 
- `batchtools` \citep{Lang2017batchtools} can [schedule jobs with Docker Swarm](https://mllg.github.io/batchtools/reference/makeClusterFunctionsDocker.html)
- scalable deployments, e.g., start with numerous Shiny talks mentioning Rocker at useR!2017
- [dynmethods](https://github.com/dynverse/dynmethods) [@rcannood]: In order to evaluate ±50 computational methods which all used different environments (R, Python, C++, ...), we wrapped each of them in a docker container and can execute these methods from R. Again, all of these containers are being built on travis-ci, and will only be pushed to docker hub if the integration test succeeds.
- \pkg{drake}, https://docs.ropensci.org/drake/index.html?q=docker#with-docker

@@

- https://github.com/benmarwick/rrtools
- \pkg{renv}, https://rstudio.github.io/renv/articles/docker.html

@@

DN conceived of the presented idea and \href{https://github.com/nuest/rockerverse-paper/issues/3}{initialised the formation of the writing team}, wrote sections not mentioned below, and revised the full first draft.
CB ..
RC and EH wrote the section on interfaces for Docker in R.
DE wrote the introduction and section about Rocker.
ME ..
CF wrote the paragraphs about `r-online`, `dockerfiler`, `r-ci` and `r-db`.
BW ..
KR ..
NR ..
NX wrote the section on liftr.
LS \& NT wrote the section on Bioconductor.
All authors contributed minor edits, gave feedback on other sections, and approved the final version.
